---
title: "Market_basket_analysis"
author: "Mohammed_Nihal"
date: "3/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Loading librarys

```{r message=FALSE}
library(RMySQL)
library(DBI)
library(arules)
library(Matrix)
library(dplyr)
library(arulesViz)
library(ggplot2)
```


# Connecting to local data base

```{r}
mydb <-  dbConnect(RMySQL::MySQL(),
                  username = "root",
                  password = "rootpassword",
                  host = "127.0.0.1",
                  port = 3306,
                  dbname = "imarket_sql"
)
```




# Extracting dataset from the local data base

```{r}
rs = dbSendQuery(mydb, "select * from line_item")
line_item = fetch(rs, n=-1)
rs = dbSendQuery(mydb, "select * from orders")
orders=fetch(rs, n=-1)
rs = dbSendQuery(mydb, "select * from products")
products=fetch(rs, n=-1)
```

### * it is seen that there are some comma's instead of decimal points in the data set

# Replace the  comma's with deimal points in unit price and total paid

```{r warning=FALSE}
line_item$unit_price=as.numeric(gsub(",",".",line_item$unit_price))
orders$total_paid=as.numeric(gsub(",",".",orders$total_paid))


```



# Merge the line_item and order

```{r}
line_order=merge.data.frame(line_item,orders,by="id_order")

```
#### we have to do some data quality process in the data set
#### first of all we need to remove incompleted orders from the data.
#### then we need to check whether is there any diffrence in the prices 
#### if so try to modify or filter the uneven changes



# Exclude not completed orders from the data set

```{r}
line_order=line_order %>% 
  filter(state=='Completed')
```


# Find the difference between the prices in line_item and order

```{r}
data_diff =line_order %>% 
  group_by(id_order) %>% 
  summarise(sum_products=sum(unit_price*product_quantity), total_paid = mean(total_paid)) %>% 
  mutate(differance =(sum_products - total_paid))
```

# Plot the graph of sum of product and total paid

```{r}
data_diff %>% ggplot(aes(sum_products,total_paid))+geom_point()+geom_abline()
```



* remove the unexplainable differences

# Set some conditions to remove the unjustifiable diffrence

```{r}
select_id_order = data_diff %>% filter(differance<=sum_products*.7,differance>=-sum_products*.6)
```


# Plot the graph after filtration

```{r}
select_id_order %>% ggplot(aes(sum_products,total_paid))+geom_point()+geom_abline()
```

# Merge product to line item

```{r}
line_product=merge(line_item,products,by="sku")
```

# Merge line product data set with the data set filtered previously

```{r}
filterd_data= merge(line_product,select_id_order,by='id_order')
filterd_data$price=as.numeric(filterd_data$price)
filterd_data=filterd_data %>% mutate(diffrence=abs(price-unit_price))
```

#### it is seen that there are some missing decimal points in the price table


# Fix the typing error in the price table

```{r}
data_error=filterd_data %>% select(id_order,name_en,sku,unit_price,price) %>%
  mutate(diffrence=abs(as.numeric(price)-unit_price)) %>% filter(diffrence>=unit_price,diffrence>=(as.numeric(price)*.95))
remaining=filterd_data %>% select(id_order,name_en,sku,unit_price,price) %>%
  mutate(diffrence=abs(as.numeric(price)-unit_price)) %>% 
  filter(!(diffrence>=unit_price & diffrence>=(as.numeric(price)*.95)))


```


# Do a for loop to find the exact position of decimal point  

```{r}

for (row in 1:nrow(data_error)) {
  price <- as.numeric(data_error$price)
  unit_price  <-as.numeric( data_error$unit_price)
  x<-nchar(as.character(as.integer(unit_price)))
  y<-nchar(as.character(as.integer(price)))
  data_error$price=(price/10^(y-x))
  data_error$diffrence=abs(price-unit_price)
 
}
 
```

# Add the modified price to the remaining data set

```{r}
filterd_data1=rbind(data_error,remaining)
```


# Remove the unjustifiable price difference

```{r}
data_diff2=filterd_data1 %>%select(id_order,name_en,unit_price,price)%>%
  mutate(diffrence=abs(as.numeric(price)-unit_price)) %>% filter(diffrence<=.75*as.numeric(price))
```


# Write a csv file

```{r}
write.csv(data_diff2,file ="filterd_dataset" )
```


# Read the dataset with  read transaction

```{r}
modal_set<-read.transactions(file = "filterd_dataset",format = c("single"),
                               header=TRUE,sep = ",",cols = c("id_order","name_en"))
```


# Find the number of items per transaction 

```{r}
table(size(modal_set))

```



# Apply the apriori rule to the transaction

```{r}
RulesName <- apriori (modal_set, parameter = list(supp = 0.0004 , conf = 0.0002257014,minlen=2))
```


# Select top ten rules and plot the graph


```{r}
top10subRules <- head(RulesName, n = 10, by = "confidence")

plot(top10subRules
     , method = "graph",  engine = "htmlwidget")
```





# top recommendations from the analysis

### * OWC Kit of 5 tools for Mac and PC  =>  Samsung 850 EVO SSD Disk 500GB
### * Smart Apple iPad Keyboard Case Keyboard Pro 10.5  <=> Apple Pencil
### * OEM 5W USB Power Adapter iPhone iPod and compatible Apple iPod <=> Apple Lightning Cable Connector to USB 1m White (OEM)
### * Spanish Keyboard Magic Apple Mac=> Apple Magic Mouse 2 <= Apple Magic Spanish Keyboard Keypad Bluetooth Mac  
### * OWC In-line Digital Temperature Sensor Kit HDD iMac 2011   <=> AdaptaDrive NewerTech adapter 2.5 to 3.5 SATA


